---
title: "Problem 3"
subtitle: "HW2"
author: "Hanyuan Chi(chixx105), Zhi Shen(shenx704)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r}
suppressPackageStartupMessages({
  library(purrr)
  library(tidyr)
  library(ggplot2)
  library(dplyr)
})

options(scipen = 99) # Penalize scientific notation

set.seed(123456) # PLEASE DO NOT CHANGE THE SEED

```

# Advanced Monte-Carlo for Poisson Process

In the context of Problem 2, please modify your solution to include that:

1. Car salesmen work 3 separate shifts: 12am-8am, 8am-4pm, 4pm-12am rather than hourly
1. Every hour of the day has its own average arrival rate
1. Car salesmen cannot work two shifts in one day.

----

You are working for Poisson Car Dealerships Inc. and your task is to optimize the employment in a particular popular car dealership location.

The customers arrive to that dealership according to a Poisson arrival process
(meaning that the number of customers that will arrive to the dealership on a particular hour is distributed with Poisson distribution and average arrival rate $\lambda$)

Salesmen are assigned to customers on 1-on-1 basis for the entire hour. Say, if 1 customer comes in at 3pm, then he occupes 1 salesman for that entire hour. The occupied salesman will be free again at 4pm to work with another customer.

Assume your dealership is open 24 hours a day / 7 days a week.

Also, assume that car salesmen work not on hourly basis but on a shift-basis. More precisely, they have 3 separate shifts: 12am-8am, 8am-4pm, 4pm-12am.
Also note that current laws prohibit the person from working more than 1 shift in a day. If a salesman works in the morning shift then he is done for the day. (The laws do permit to work adjacent shift across different days: say, it is allowed to have 4pm-12am shift one day followed immediately by 12am-8am shift the next day.)

----

## Question 1

* Please simulate one possible future for the next month (30 days = 720 hours).

* Output:
    - Please create data.frame `df1` with $N$ rows that contains your sample values in column `df1$X`, current hour (in 0-23 format) in `df1$hour` and current day (in 1-30 format) `df1$day`.

```{r}
N <- 720L

# Specify a given average arrival rate for every hour
AVG_ARRIVAL_RATE <- 10*sin(seq(0,3*pi/4,length.out = 24))
names(AVG_ARRIVAL_RATE) <- 0:23 # 24-hour based time

# Print the (rounded) arrival rate for every hour of the day
print(round(AVG_ARRIVAL_RATE,2))

# Please write your code below

df1 <- data.frame(hour = rep(0:23,N/24),
                  day = rep(1:30, each =24))%>%
  group_by(hour)%>%
  mutate(X = rpois(30,AVG_ARRIVAL_RATE[hour+1]))

head(df1)
```

## Question 2

* Please simulate $R=2000$ possible ways your future may look for the next month (30 days = 720 hours)


* Output:
    - Please create data.frame `df2` with $N\times R$ rows that contains your sample values in column `df2$X`, current hour (in 0-23 format) in `df2$hour`, current day (in 1-30 format) `df2$day` and sample id in column `df2$id`.

```{r}
R <- 2000L

# Please write your code below

df2 <- data.frame(id = rep(1:R, each=N),
                  hour = rep(0:23, R*N/24),
                  day = rep(rep(1:30, each=24),R))%>%
  group_by(hour)%>%
  mutate(X = rpois(30*R,AVG_ARRIVAL_RATE[hour+1]))

head(df2)
```

## Question 3

* Please compute how much workforce you should deploy every day to make sure that in **99% of the days** there are enough salesmen present for every customer (without waiting).

* Hints:
    - you may want to look at `quantile()` function

* Output:
    - Please save the value into integer variable `q3` 

```{r}
# Please write your code below

df3 <- df2 %>%
  mutate(shift = ifelse(hour %in% c(0:7),1,ifelse(hour %in% c(8:15),2,3))) %>%
  group_by(id,day,shift) %>%
  summarise(max_shift_X = max(X))%>%
  group_by(id,day) %>%
  summarise(day_X = sum(max_shift_X))
         
q3 <- as.integer(quantile(df3$day_X,0.99))
q3
```


