---
title: "Problem 2"
subtitle: "HW2"
author: "Hanyuan Chi(chixx105), Zhi Shen(shenx704)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r}
suppressPackageStartupMessages({
  library(purrr)
  library(tidyr)
  library(ggplot2)
  library(dplyr)
})

set.seed(123456) # PLEASE DO NOT CHANGE THE SEED

```

# Monte-Carlo Simulations for Poisson Process

You are working for Poisson Car Dealerships Inc. and your task is to optimize the employment in a particular popular car dealership location.

The customers arrive to that dealership according to a Poisson arrival process
(meaning that the number of customers that will arrive to the dealership on a particular hour is distributed with Poisson distribution and average arrival rate $\lambda$)

Salesmen are assigned to customers on 1-on-1 basis. Say, if 1 customer comes in at 3pm, then he occupies 1 salesman for that entire hour. The occupied salesman will be free again at 4pm to work with another customer.

Assume your dealership is open 24 hours a day / 7 days a week.

## Question 1

* Please simulate one possible future for the next month (30 days = 720 hours) assuming the average arrival rate $\lambda=6$ customers per hour.

* Hints:
    - you will need to use what you know about Poisson distribution


* Output:
    - Please create data.frame `df1` with $N$ rows that contains your sample values in column `df1$X`, current hour (in 0-23 format) in `df1$hour` and current day (in 1-30 format) `df1$day`.

```{r}
pois_lambda <- 6
N <- 720L

# Please write your code below

df1 <- data.frame(X = rpois(N,pois_lambda),
                  hour = rep(0:23, N/24),
                  day = rep(1:30, each=24))
head(df1)
```

## Question 2

* Please simulate $R=2000L$ possible ways your future may look for the next month (30 days = 720 hours)


* Output:
    - Please create data.frame `df2` with $N\times R$ rows that contains your sample values in column `df2$X`, current hour (in 0-23 format) in `df2$hour`, current day (in 1-30 format) `df2$day` and sample id in column `df2$id`.

```{r}
R <- 2000L

# Please write your code below

df2 <- data.frame(id = rep(1:R, each=N),
                  X = rpois(R*N,pois_lambda),
                  hour = rep(0:23, R*N/24),
                  day = rep(rep(1:30, each=24),R))
head(df2)

```

## Question 3

* Please compute how many salesmen you should keep on duty each hour to make sure that in 99% of **hours** there are enough salesmen for every customer (without waiting).

* Hints:
    - you may want to look at `quantile()` function

* Output:
    - Please save the value into numeric variable `q3`

```{r}
# Please write your code below

q3 <- quantile(df2$X,0.99)
q3
```
